A novel catalyitic model for estimating age-specific measles sero-prevalence uisng case-incidence data: a case study from the Democratic Republic of the Congo

### 2018-06-22 Paper/paper.tex
efficacy --> effectiveness

\subsection{Simulated Data\label{subsec:SimData}}
To confirm that we can accurately recover the force of infection and vaccination hazard using the model we have chosen, we generate simulated datasets consisting of age-specified measles case, vaccination, and serology data using a previously developed, age-structured MSIRV (Maternally immune, Susceptible, Infected, Recovered, Vaccinated) model~\cite{Metcalf2012}. We used demographic parameters from UN estimates for the DRC and increased vaccination linearly from 0 to 50\% over the first 30 years of a 50 year simulation. $R_{0}$ (the number of secondary cases resulting from the introduction of a single infected individual) was assumed to be constant at 15 over the entirety of the simulation. The force of infection (foi) was not directly chosen to be a Wiebull function, but is generated using the specified $R_0$ and WAIFW (Who Acquires Infection from Whom) matrix that describes social interactions as estimated via the POLYMOD study~\cite{Polymod}. This means that we cannot directly compare the recovered Weibull parameters to injected parameters, but we \emph{can} compare the recovered foi curve to the true values, as well as the recovered vaccination efficacy ($\gamma$) to the true value.

@@

\subsection{Comparing Data and Inference \label{subsec:Comp}}
A standard approach for interpreting the results of Bayesian inference would be inspecting the posterior distributions for the model parameters. But the parameters of our model are not, in themselves, particularly interesting - it is the inferred seroprevalence, case distribution, and vaccination probability that we wish to compare with the data. To do this, we generate the posteriors on $\{\eta_f, \alpha_f, \beta_f, \gamma, \eta_v, \alpha_v, \beta_v\}$ as described, and then draw from these posteriors to generate the seroprevalence, case, and vaccination curves as laid out in Sec.~\ref{subsec:CompetingRates}. Because the vaccine efficacy, $\gamma$, is of interest on its own, we also examine the posterior distributions for this parameter in detail.

@@

The vaccine effectiveness, $\gamma$, which measures the likelihood that a child recorded as vaccinated in a vaccine survey was immunized by vaccine, is poorly estimated using a model analyzed with only age-specific case incidence and vaccination survey data (Figure~\ref{fig:simplots} (c) - top).  The inferred values of vaccine efficacy are positively correlated with the true value used in simulation, but are biased significantly low; in simulation runs with low effectiveness, the resulting posterior estimate was low and significantly different than the prior on $\gamma$ (Figure~\ref{fig:simplots} f).  This bias is off-set by slight biases in other parameters (see Supplement), which explains why the estimated age-specific sero-prevalence only exhibited a slight negative bias. Estimated vaccine effectiveness remains biased in the corresponding model that includes seroprevalence (Eq 8); though the correlation between the posterior mean estimate and the true value used in simulation is stronger and the variance of the posterior distribution is smaller when serology data are included (Figure~\ref{fig:simplots} (c) - bottom).  Thus, while vaccine effectiveness may always be challenging to estimate, the resulting estimate from this model may serve as an indicator of vaccine effectiveness, and in particular the \emph{relative} values of vaccine effectiveness recovered for two different locations likely correctly identify the location with lower $\gamma$.  Results for the age-specific sero-prevalence, probability of vaccination, and the hazards for FOI and vaccination (e.g. corresponding to (Figure~\ref{fig:simplots}) for the model analyzed using age-specific case incidence, vaccination survey data, and seroprevalence data are presented in the Supplement.

@@

\begin{figure*}
\includegraphics[width=\textwidth,angle=0]{figures/VacFOIveff-crop.pdf}
\caption{\label{fig:foivaccveff} Panel (a): Vaccination hazard (solid lines, right y-axis) and vaccination coverage (dashed lines, left y-axis) recovered using all three sets of data from all 11 provinces, generated from a set of posterior samples. For most provinces, the vaccination hazard is tightly peaked near one year, as expected. Panel (b): The recovered force of infection using all data (solid curve) and only vaccination and case data (dashed curves). As in Figure~\ref{fig:sero}, the solid curve is the mean of the posterior distribution, and the provinces before the blank space were fit using the Weibull plus constant model while those after were fit using the constant-only model for force of infection. In most cases, the inclusion of the serology data did not make a large difference in the inferred force of infection. Panel (c): recovered value for vaccine efficacy ($\pm 1\sigma$) from analyses using all three data sets for each province (solid), and vaccine and case data only (dashed). Note that the values without serology data are almost all consistent with the prior on $\gamma$. Because of our results from the simulated data, we do not expect that these values to reflect the true vaccine efficacy (in general, they are probably biased low), but we \emph{do} expect that the ranking of provinces by vaccine efficacy is correct.}
\end{figure*}

>>>>

\subsection{Simulated Data\label{subsec:SimData}}
To confirm that we can accurately recover the force of infection and vaccination hazard using the model we have chosen, we generate simulated datasets consisting of age-specified measles case, vaccination, and serology data using a previously developed, age-structured MSIRV (Maternally immune, Susceptible, Infected, Recovered, Vaccinated) model~\cite{Metcalf2012}. We used demographic parameters from UN estimates for the DRC and increased vaccination linearly from 0 to 50\% over the first 30 years of a 50 year simulation. $R_{0}$ (the number of secondary cases resulting from the introduction of a single infected individual) was assumed to be constant at 15 over the entirety of the simulation. The force of infection (foi) was not directly chosen to be a Wiebull function, but is generated using the specified $R_0$ and WAIFW (Who Acquires Infection from Whom) matrix that describes social interactions as estimated via the POLYMOD study~\cite{Polymod}. This means that we cannot directly compare the recovered Weibull parameters to injected parameters, but we \emph{can} compare the recovered foi curve to the true values, as well as the recovered vaccination effectiveness ($\gamma$) to the true value.

@@

\subsection{Comparing Data and Inference \label{subsec:Comp}}
A standard approach for interpreting the results of Bayesian inference would be inspecting the posterior distributions for the model parameters. But the parameters of our model are not, in themselves, particularly interesting - it is the inferred seroprevalence, case distribution, and vaccination probability that we wish to compare with the data. To do this, we generate the posteriors on $\{\eta_f, \alpha_f, \beta_f, \gamma, \eta_v, \alpha_v, \beta_v\}$ as described, and then draw from these posteriors to generate the seroprevalence, case, and vaccination curves as laid out in Sec.~\ref{subsec:CompetingRates}. Because the vaccination effectiveness, $\gamma$, is of interest on its own, we also examine the posterior distributions for this parameter in detail.

@@

The vaccine effectiveness, $\gamma$, which measures the likelihood that a child recorded as vaccinated in a vaccine survey was immunized by vaccine, is poorly estimated using a model analyzed with only age-specific case incidence and vaccination survey data (Figure~\ref{fig:simplots} (c) - top).  The inferred values of vaccination effictiveness are positively correlated with the true value used in simulation, but are biased significantly low; in simulation runs with low effectiveness, the resulting posterior estimate was low and significantly different than the prior on $\gamma$ (Figure~\ref{fig:simplots} f).  This bias is off-set by slight biases in other parameters (see Supplement), which explains why the estimated age-specific sero-prevalence only exhibited a slight negative bias. Estimated vaccine effectiveness remains biased in the corresponding model that includes seroprevalence (Eq 8); though the correlation between the posterior mean estimate and the true value used in simulation is stronger and the variance of the posterior distribution is smaller when serology data are included (Figure~\ref{fig:simplots} (c) - bottom).  Thus, while vaccine effectiveness may always be challenging to estimate, the resulting estimate from this model may serve as an indicator of vaccine effectiveness, and in particular the \emph{relative} values of vaccine effectiveness recovered for two different locations likely correctly identify the location with lower $\gamma$.  Results for the age-specific sero-prevalence, probability of vaccination, and the hazards for FOI and vaccination (e.g. corresponding to (Figure~\ref{fig:simplots}) for the model analyzed using age-specific case incidence, vaccination survey data, and seroprevalence data are presented in the Supplement.

@@

\begin{figure*}
\includegraphics[width=\textwidth,angle=0]{figures/VacFOIveff-crop.pdf}
\caption{\label{fig:foivaccveff} Panel (a): Vaccination hazard (solid lines, right y-axis) and vaccination coverage (dashed lines, left y-axis) recovered using all three sets of data from all 11 provinces, generated from a set of posterior samples. For most provinces, the vaccination hazard is tightly peaked near one year, as expected. Panel (b): The recovered force of infection using all data (solid curve) and only vaccination and case data (dashed curves). As in Figure~\ref{fig:sero}, the solid curve is the mean of the posterior distribution, and the provinces before the blank space were fit using the Weibull plus constant model while those after were fit using the constant-only model for force of infection. In most cases, the inclusion of the serology data did not make a large difference in the inferred force of infection. Panel (c): recovered value for vaccination effictiveness ($\pm 1\sigma$) from analyses using all three data sets for each province (solid), and vaccine and case data only (dashed). Note that the values without serology data are almost all consistent with the prior on $\gamma$. Because of our results from the simulated data, we do not expect that these values to reflect the true vaccination effectiveness (in general, they are probably biased low), but we \emph{do} expect that the ranking of provinces by vaccination effectiveness is correct.}
\end{figure*}
